---
title: "Getting Started with pluotr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with pluotr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

pluotr is an R interface to [pluot](https://pluot.dev), a GPU-accelerated
rendering library written in Rust. You describe a plot as a list of layers and
pluot produces either raster (RGBA pixel bytes) or vector (SVG) output.

## Installation

```{r, eval=FALSE}
# From the pluot workspace root (requires Rust/Cargo)
# On macOS also install LLVM: brew install llvm
devtools::install("bindings-r")
```

## A helper for displaying raster output

`pluot_render()` returns a raw vector of `width * height * 4` RGBA bytes plus
one trailing status byte. The helper below converts it into a base-R raster
object that `plot()` can display directly.

```{r raster-helper}
library(pluotr)

pluot_raster <- function(raw_bytes, width, height) {
  pixel_bytes <- raw_bytes[-length(raw_bytes)]          # drop status byte
  vals <- as.double(as.integer(pixel_bytes)) / 255.0
  arr  <- array(vals, dim = c(4L, width, height))       # [channel, x, y]
  as.raster(aperm(arr, c(3L, 2L, 1L)))                  # --> [y, x, channel]
}
```

## Raster output

The `camera_view` is a column-major 4×4 view matrix. An identity matrix maps
pixel coordinates directly to the canvas without any scaling or translation.

Layers are named lists with `layer_type` (a string matching a registered layer)
and `layer_params` (layer-specific settings). `PointLayer` accepts inline
coordinate vectors in either pixel or data space.

```{r raster, fig.width=3, fig.height=3}
camera_identity <- c(
  1, 0, 0, 0,
  0, 1, 0, 0,
  0, 0, 1, 0,
  0, 0, 0, 1
)

result <- pluot_render(
  layers = list(
    list(
      layer_type   = "PointLayer",
      layer_params = list(
        layer_id                 = "four_points",
        data_unit_mode_x         = "Pixels",
        data_unit_mode_y         = "Pixels",
        point_radius_unit_mode_x = "Pixels",
        point_radius_unit_mode_y = "Pixels",
        point_shape_mode         = "Circle",
        point_radius             = 20.0,
        bounds                   = NULL,
        position_x               = c(60,  60, 140, 140),
        position_y               = c(60, 140,  60, 140),
        labels_vec               = c(0L, 1L, 2L, 3L)
      )
    )
  ),
  width       = 200L,
  height      = 200L,
  camera_view = camera_identity,
  plot_id     = "intro_raster"
)

plot(pluot_raster(result, 200L, 200L), interpolate = FALSE)
```

## SVG output

Pass `format = "Vector"` to receive a UTF-8 SVG string instead of pixel bytes.
This is useful for publication figures or when you want to embed the chart in an
HTML document.

```{r svg}
svg_result <- pluot_render(
  layers = list(
    list(
      layer_type   = "PointLayer",
      layer_params = list(
        layer_id                 = "four_points",
        data_unit_mode_x         = "Pixels",
        data_unit_mode_y         = "Pixels",
        point_radius_unit_mode_x = "Pixels",
        point_radius_unit_mode_y = "Pixels",
        point_shape_mode         = "Circle",
        point_radius             = 20.0,
        bounds                   = NULL,
        position_x               = c(60,  60, 140, 140),
        position_y               = c(60, 140,  60, 140),
        labels_vec               = c(0L, 1L, 2L, 3L)
      )
    )
  ),
  width       = 200L,
  height      = 200L,
  format      = "Vector",
  camera_view = camera_identity,
  plot_id     = "intro_svg"
)

htmltools::HTML(rawToChar(svg_result))
```
