---
title: "Layer Types"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Layer Types}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r helpers, include=FALSE}
library(pluotr)

camera_identity <- c(
  1, 0, 0, 0,
  0, 1, 0, 0,
  0, 0, 1, 0,
  0, 0, 0, 1
)
```

pluot builds plots by compositing layers. Each layer is a named list with a
`layer_type` string and a `layer_params` list. This vignette walks through the
built-in layer types.

## PointLayer

`PointLayer` renders a set of points at inline pixel or data coordinates.
Key parameters:

| Parameter | Description |
|-----------|-------------|
| `position_x`, `position_y` | Coordinates as `list(dtype = ..., values = ...)`, where `dtype` is one of `"Uint8"`, `"Uint16"`, `"Uint32"`, `"Uint64"`, `"Int8"`, `"Int16"`, `"Int32"`, `"Int64"`, `"Float32"`, `"Float64"` and `values` is a numeric vector |
| `labels_vec` | Integer vector of group indices (drives colour) |
| `point_radius` | Radius in the units given by `point_radius_unit_mode_*` |
| `point_shape_mode` | `"Circle"` or `"Square"` |
| `data_unit_mode_*` | `"Pixels"` or `"Data"` |

```{r point-layer, fig.width=3, fig.height=3}
display_raster(
  layers = list(
    list(
      layer_type   = "PointLayer",
      layer_params = list(
        layer_id                 = "pts",
        data_unit_mode_x         = "Pixels",
        data_unit_mode_y         = "Pixels",
        point_radius_unit_mode_x = "Pixels",
        point_radius_unit_mode_y = "Pixels",
        point_shape_mode         = "Circle",
        point_radius             = 18.0,
        bounds                   = NULL,
        position_x               = list(dtype = "Float32", values = c(50, 100, 150, 50, 150)),
        position_y               = list(dtype = "Float32", values = c(50,  50,  50, 150, 150)),
        labels_vec               = c(0L, 1L, 2L, 3L, 4L)
      )
    )
  ),
  width       = 200L,
  height      = 200L,
  camera_view = camera_identity,
  plot_id     = "lt_point"
)
```

## AxisLinearLayer

`AxisLinearLayer` draws a linear axis along one edge of the drawable area.
Set `position` to `"Left"`, `"Right"`, `"Top"`, or `"Bottom"`. Axes require
matching margins so they have space to render outside the data area.

```{r axis-layer, fig.width=4, fig.height=4}
display_raster(
  layers = list(
    list(
      layer_type   = "PointLayer",
      layer_params = list(
        layer_id                 = "pts",
        data_unit_mode_x         = "Pixels",
        data_unit_mode_y         = "Pixels",
        point_radius_unit_mode_x = "Pixels",
        point_radius_unit_mode_y = "Pixels",
        point_shape_mode         = "Circle",
        point_radius             = 8.0,
        bounds                   = NULL,
        position_x               = list(dtype = "Float32", values = c(100, 160, 220, 280, 340)),
        position_y               = list(dtype = "Float32", values = c(200, 140,  80, 200, 120)),
        labels_vec               = c(0L, 1L, 2L, 3L, 4L)
      )
    ),
    list(
      layer_type   = "AxisLinearLayer",
      layer_params = list(layer_id = "left_axis",   position = "Left")
    ),
    list(
      layer_type   = "AxisLinearLayer",
      layer_params = list(layer_id = "bottom_axis", position = "Bottom")
    )
  ),
  width         = 400L,
  height        = 400L,
  camera_view   = camera_identity,
  margin_left   = 40L,
  margin_bottom = 40L,
  margin_top    = 20L,
  margin_right  = 20L,
  plot_id       = "lt_axis"
)
```

## TextLayer

`TextLayer` places text labels at arbitrary positions. Like `PointLayer` it
accepts `data_unit_mode_*` (`"Pixels"` or `"Data"`). `text_rotation` rotates
the label in degrees (anti-clockwise).

```{r text-layer}
w <- 400L
h <- 300L
ml <- 70L; mb <- 60L; mt <- 30L; mr <- 20L

display_svg(
  layers = list(
    list(
      layer_type   = "PointLayer",
      layer_params = list(
        layer_id                 = "pts",
        data_unit_mode_x         = "Pixels",
        data_unit_mode_y         = "Pixels",
        point_radius_unit_mode_x = "Pixels",
        point_radius_unit_mode_y = "Pixels",
        point_shape_mode         = "Circle",
        point_radius             = 8.0,
        bounds                   = NULL,
        position_x               = list(dtype = "Float32", values = c(100, 160, 220, 280, 340)),
        position_y               = list(dtype = "Float32", values = c(180, 130,  80, 200, 110)),
        labels_vec               = c(0L, 1L, 2L, 3L, 4L)
      )
    ),
    list(
      layer_type   = "AxisLinearLayer",
      layer_params = list(layer_id = "left_axis",   position = "Left")
    ),
    list(
      layer_type   = "AxisLinearLayer",
      layer_params = list(layer_id = "bottom_axis", position = "Bottom")
    ),
    # Horizontal axis title centred below the bottom axis
    list(
      layer_type   = "TextLayer",
      layer_params = list(
        layer_id             = "x_title",
        data_unit_mode_x     = "Pixels",
        data_unit_mode_y     = "Pixels",
        text_size_unit_mode  = "Pixels",
        text_size            = 14.0,
        text_align_mode      = "Middle",
        text_baseline_mode   = "Middle",
        font_family = NULL,
        font_weight = "Normal",
        font_style = "Normal",
        bounds               = list(
          margin_top = 0, margin_left = 0,
          margin_right = 0, margin_bottom = 0
        ),
        position_x           = list(dtype = "Float32", values = list(ml + (w - ml - mr) / 2)),
        position_y           = list(dtype = "Float32", values = list(h - 12.0)),
        text_vec             = list("X axis")
      )
    ),
    # Vertical axis title rotated 90° along the left margin
    list(
      layer_type   = "TextLayer",
      layer_params = list(
        layer_id             = "y_title",
        data_unit_mode_x     = "Pixels",
        data_unit_mode_y     = "Pixels",
        text_size_unit_mode  = "Pixels",
        text_size            = 14.0,
        text_align_mode      = "Middle",
        text_baseline_mode   = "Middle",
        text_rotation        = -90.0,
        font_family = NULL,
        font_weight = "Normal",
        font_style = "Normal",
        bounds               = list(
          margin_top = 0, margin_left = 0,
          margin_right = 0, margin_bottom = 0
        ),
        position_x           = list(dtype = "Float32", values = list(18.0)),
        position_y           = list(dtype = "Float32", values = list(mt + (h - mt - mb) / 2)),
        text_vec             = list("Y axis")
      )
    )
  ),
  width         = w,
  height        = h,
  camera_view   = camera_identity,
  margin_left   = ml,
  margin_bottom = mb,
  margin_top    = mt,
  margin_right  = mr,
  plot_id       = "lt_text"
)
```

## Compositing layers

Layers are drawn in list order (first = bottom, last = top). You can overlay
any combination of the types above to build complex plots.

```{r composite}
set.seed(7)
n  <- 200L
xs <- 80 + runif(n) * 240
ys <- 40 + runif(n) * 180

display_svg(
  layers = list(
    list(
      layer_type   = "PointLayer",
      layer_params = list(
        layer_id                 = "bg",
        data_unit_mode_x         = "Pixels",
        data_unit_mode_y         = "Pixels",
        point_radius_unit_mode_x = "Pixels",
        point_radius_unit_mode_y = "Pixels",
        point_shape_mode         = "Circle",
        point_radius             = 4.0,
        bounds                   = NULL,
        position_x               = list(dtype = "Float32", values = xs),
        position_y               = list(dtype = "Float32", values = ys),
        labels_vec               = as.integer(rep_len(0L:4L, n))
      )
    ),
    list(
      layer_type   = "AxisLinearLayer",
      layer_params = list(layer_id = "left",   position = "Left")
    ),
    list(
      layer_type   = "AxisLinearLayer",
      layer_params = list(layer_id = "bottom", position = "Bottom")
    ),
    list(
      layer_type   = "TextLayer",
      layer_params = list(
        layer_id            = "title",
        data_unit_mode_x    = "Pixels",
        data_unit_mode_y    = "Pixels",
        text_size_unit_mode = "Pixels",
        text_size           = 16.0,
        text_align_mode     = "Middle",
        text_baseline_mode  = "Middle",
        font_family = NULL,
        font_weight = "Normal",
        font_style = "Normal",
        bounds              = list(margin_top=0, margin_left=0,
                                   margin_right=0, margin_bottom=0),
        position_x          = list(dtype = "Float32", values = list(200.0)),
        position_y          = list(dtype = "Float32", values = list(14.0)),
        text_vec            = list("Composite plot")
      )
    )
  ),
  width         = 400L,
  height        = 300L,
  camera_view   = camera_identity,
  margin_left   = 70L,
  margin_bottom = 50L,
  margin_top    = 30L,
  margin_right  = 20L,
  plot_id       = "lt_composite"
)
```
