---
title: "Scatter Plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scatter Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r helpers, include=FALSE}
library(pluotr)
```

This vignette shows how to build scatter plots with axes, margins, and a
calibrated camera view.

## Data in pixel space

The simplest approach is to scale your data to pixel coordinates directly and
use the identity camera. Here 500 random points are placed within the inner
area of a 500×500 canvas, leaving room for axis labels on the left and bottom.

```{r scatter-pixels, fig.width=5, fig.height=5}
set.seed(42)
n      <- 500L
margin <- 80L          # pixels reserved for axis labels
w      <- 500L
h      <- 500L

# Scale data to the drawable area [margin, w-20] × [20, h-margin]
x_px <- runif(n) * (w - margin - 20L)
y_px <- runif(n) * (h - margin - 20L)

# Assign 5 categorical groups (0-based integer labels drive colour)
groups <- as.integer(rep_len(0L:4L, n))

camera_identity <- c(
  1, 0, 0, 0,
  0, 1, 0, 0,
  0, 0, 1, 0,
  0, 0, 0, 1
)

display_raster(
  layers = list(
    list(
      layer_type   = "PointLayer",
      layer_params = list(
        layer_id                 = "scatter",
        data_unit_mode_x         = "Pixels",
        data_unit_mode_y         = "Pixels",
        point_radius_unit_mode_x = "Pixels",
        point_radius_unit_mode_y = "Pixels",
        point_shape_mode         = "Circle",
        point_radius             = 5.0,
        bounds                   = NULL,
        position_x               = list(dtype = "Float32", values = x_px),
        position_y               = list(dtype = "Float32", values = y_px),
        labels_vec               = groups
      )
    ),
    list(
      layer_type   = "AxisLinearLayer",
      layer_params = list(layer_id = "left_axis",   position = "Left")
    ),
    list(
      layer_type   = "AxisLinearLayer",
      layer_params = list(layer_id = "bottom_axis", position = "Bottom")
    )
  ),
  width         = w,
  height        = h,
  camera_view   = camera_identity,
  margin_left   = margin,
  margin_bottom = margin,
  margin_top    = 20L,
  margin_right  = 20L,
  plot_id       = "scatter_pixels"
)
```

## SVG output for publication

Replace `format = "Raster"` with `format = "Vector"` to get SVG. The same
layer definition works for both output modes.

```{r scatter-svg}
display_svg(
  layers = list(
    list(
      layer_type   = "PointLayer",
      layer_params = list(
        layer_id                 = "scatter",
        data_unit_mode_x         = "Pixels",
        data_unit_mode_y         = "Pixels",
        point_radius_unit_mode_x = "Pixels",
        point_radius_unit_mode_y = "Pixels",
        point_shape_mode         = "Circle",
        point_radius             = 5.0,
        bounds                   = NULL,
        position_x               = list(dtype = "Float32", values = x_px),
        position_y               = list(dtype = "Float32", values = y_px),
        labels_vec               = groups
      )
    ),
    list(
      layer_type   = "AxisLinearLayer",
      layer_params = list(layer_id = "left_axis",   position = "Left")
    ),
    list(
      layer_type   = "AxisLinearLayer",
      layer_params = list(layer_id = "bottom_axis", position = "Bottom")
    )
  ),
  width         = w,
  height        = h,
  camera_view   = camera_identity,
  margin_left   = margin,
  margin_bottom = margin,
  margin_top    = 20L,
  margin_right  = 20L,
  plot_id       = "scatter_svg"
)
```

