---
title: "Remote Zarr Stores over HTTP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Remote Zarr Stores over HTTP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r helpers, include=FALSE}
library(pluotr)

pluot_raster <- function(raw_bytes, width, height) {
  pixel_bytes <- raw_bytes[-length(raw_bytes)]
  vals <- as.double(as.integer(pixel_bytes)) / 255.0
  arr  <- array(vals, dim = c(4L, width, height))
  as.raster(aperm(arr, c(3L, 2L, 1L)))
}
```

pluotr can render data from any Zarr store that pizzarr supports, including
stores served over HTTP. This vignette demonstrates `ZarrBarPlotLayer` using a
publicly available dataset — historical UK wheat prices — fetched directly from
a remote Zarr store.

> **Note:** this vignette requires an internet connection and the
> [`crul`](https://docs.ropensci.org/crul/) package:
> ```r
> install.packages("crul")
> ```

## Opening a remote Zarr store

pizzarr's `HttpStore` fetches chunks on demand over HTTP. Pass the root URL of
the Zarr store to the constructor.

```{r open-store}
library(pizzarr)

store <- HttpStore$new(
  "https://pub-adb3658c8ed642caa534fdc612cd1c0c.r2.dev/wheat.zarr"
)
store
```

## Registering the store with pluotr

Give the store a name and hand it to `pluot_register_store()`. Use the same
name as `store_name` in `pluot_render()`.

```{r register-store}
library(pluotr)

pluot_register_store("wheat_store", store)
```

## Rendering a bar plot

`ZarrBarPlotLayer` reads two arrays from the store:

| parameter | array | description |
|---|---|---|
| `identifier_key` | `/year` | category axis (bar labels) |
| `quantity_key` | `/wheat` | value axis (bar heights) |

The camera view matrix controls which part of the data space is visible. Here
the x-scale (`0.15`) spreads out the bars and the y-scale (`0.01`) compresses
the tall price axis into the canvas height. The y-translation (`-1.0` in the
last column) shifts the origin to the bottom of the plot.

```{r render-raster, fig.width=8, fig.height=6}
w <- 800L
h <- 600L

camera_view <- c(
  0.15, 0.0,  0.0, 0.0,
  0.0,  0.01, 0.0, 0.0,
  0.0,  0.0,  1.0, 0.0,
  0.0,  -1.0, 0.0, 1.0
)

result <- pluot_render(
  layers = list(
    list(
      layer_type   = "ZarrBarPlotLayer",
      layer_params = list(
        layer_id        = "barplot_layer",
        bounds          = NULL,
        orientation     = "Vertical",
        identifier_key  = "/year",
        quantity_key    = "/wheat",
        fill_color_mode = "Categorical"
      )
    )
  ),
  width               = w,
  height              = h,
  camera_view         = camera_view,
  margin_left         = 100L,
  margin_bottom       = 100L,
  margin_top          = 10L,
  margin_right        = 10L,
  plot_id             = "wheat_barplot",
  store_name          = "wheat_store",
  wait_for_store_gets = TRUE,
  cache_enabled       = FALSE
)

plot(pluot_raster(result, w, h), interpolate = FALSE)
```

## SVG output

The same layer definition produces publication-quality SVG by passing
`format = "Vector"`.

```{r render-svg}
svg_result <- pluot_render(
  layers = list(
    list(
      layer_type   = "ZarrBarPlotLayer",
      layer_params = list(
        layer_id        = "barplot_layer",
        bounds          = NULL,
        orientation     = "Vertical",
        identifier_key  = "/year",
        quantity_key    = "/wheat",
        fill_color_mode = "Categorical"
      )
    )
  ),
  width               = w,
  height              = h,
  format              = "Vector",
  camera_view         = camera_view,
  margin_left         = 100L,
  margin_bottom       = 100L,
  margin_top          = 10L,
  margin_right        = 10L,
  plot_id             = "wheat_barplot_svg",
  store_name          = "wheat_store",
  wait_for_store_gets = TRUE,
  cache_enabled       = FALSE
)

htmltools::HTML(rawToChar(svg_result))
```
