---
title: "Zarr-backed Scatter Plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Zarr-backed Scatter Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r helpers, include=FALSE}
library(pluotr)

pluot_raster <- function(raw_bytes, width, height) {
  pixel_bytes <- raw_bytes[-length(raw_bytes)]
  vals <- as.double(as.integer(pixel_bytes)) / 255.0
  arr  <- array(vals, dim = c(4L, width, height))
  as.raster(aperm(arr, c(3L, 2L, 1L)))
}
```

`ZarrPointLayer` reads point coordinates and colour labels directly from
[Zarr v3](https://zarr.dev) arrays. This is useful when your data already lives
in a Zarr store on disk or in the cloud — pluotr fetches only the chunks it
needs through a set of R callbacks, so the full array never has to be loaded
into memory at once.

This vignette shows how to build a Zarr v3 store in memory using
[pizzarr](https://zarr-developers.github.io/pizzarr/) and render it with
`ZarrPointLayer`.

## Prerequisites

pizzarr >= 0.2.0 is required for `zarr_format = 3L` support.
See `bindings-r/README.md` in the Pluot repository for troubleshooting instructions if needed.

```{r install-pizzarr, eval=FALSE}
install.packages("pizzarr")
```

## Building an in-memory Zarr v3 store

`zarr_create_group(zarr_format = 3L)` creates an in-memory group that stores
metadata in `zarr.json` files (Zarr v3 format). Each `create_dataset` call adds
one named array to the group.

`ZarrPointLayer` expects:
- **x / y arrays** — `dtype = "<f8"` (64-bit float)
- **colour labels array** — `dtype = "<i8"` (64-bit integer, 0-based category
  index)

```{r build-store}
library(pizzarr)

set.seed(42)
n      <- 200L
groups <- as.integer(rep_len(0L:4L, n))   # 5 colour categories

g <- zarr_create_group(zarr_format = 3L)
g$create_dataset("x",      data = array(runif(n)),         shape = n, dtype = "<f8")
g$create_dataset("y",      data = array(runif(n)),         shape = n, dtype = "<f8")
g$create_dataset("labels", data = array(as.double(groups)), shape = n, dtype = "<i8")
```

## Registering the store with pluotr

`pluot_register_store()` gives the store a name that `ZarrPointLayer` can look
up at render time. The name is an arbitrary string; pass the same value as
`store_name` in `pluot_render()`.

```{r register-store}
library(pluotr)

pluot_register_store("demo_store", g$get_store())
```

## Rendering

`ZarrPointLayer` reads coordinates in **data space** (the unit range `[0, 1]`
for both axes here) and colours each point by its label. The `x_key`, `y_key`,
and `color_key` parameters are absolute paths within the store.

Set `wait_for_store_gets = TRUE` so that Rust waits for the synchronous R zarr
callbacks before proceeding.

```{r render, fig.width=5, fig.height=5}
w <- 500L
h <- 500L

result <- pluot_render(
  layers = list(
    list(
      layer_type   = "ZarrPointLayer",
      layer_params = list(
        layer_id                 = "scatter",
        data_unit_mode_x         = "Data",
        data_unit_mode_y         = "Data",
        point_radius_unit_mode_x = "Pixels",
        point_radius_unit_mode_y = "Pixels",
        point_shape_mode         = "Circle",
        point_radius             = 6.0,
        bounds                   = NULL,
        store_name               = NULL,
        x_key                    = "/x",
        y_key                    = "/y",
        color_key                = "/labels"
      )
    )
  ),
  width               = w,
  height              = h,
  camera_view         = c(1,0,0,0, 0,1,0,0, 0,0,1,0, 0,0,0,1),
  plot_id             = "zarr_scatter",
  store_name          = "demo_store",
  wait_for_store_gets = TRUE,
  cache_enabled       = FALSE
)

plot(pluot_raster(result, w, h), interpolate = FALSE)
```

## Using a real Zarr store on disk

The same approach works for stores on disk: open the store with `pizzarr`, then
register it with `pluot_register_store()`. pluotr will call back into R to
fetch each chunk via the standard pizzarr store interface, so any store type
that pizzarr supports (local filesystem, HTTP, S3, …) works without any changes
to the rendering code.

```{r disk-store, eval=FALSE}
disk_store <- pizzarr::DirectoryStore$new("/path/to/my.zarr")
pluot_register_store("disk_store", disk_store)

pluot_render(
  layers = list(
    list(
      layer_type   = "ZarrPointLayer",
      layer_params = list(
        layer_id    = "scatter",
        # ... same params as above ...
        x_key       = "/coords/x",
        y_key       = "/coords/y",
        color_key   = "/obs/cluster"
      )
    )
  ),
  store_name          = "disk_store",
  wait_for_store_gets = TRUE,
  # ...
)
```
